---
title: "Phenotype-guided SigBridgeR pipeline"
subtitle: "Neural-biased morphotype: preprocessing, four-method screening, consensus, and visualisation"
from: markdown+emoji
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: show
    code-tools: true
    embed-resources: true
    html-math-method: mathjax
execute:
  echo: true
  warning: true
  message: true
  error: false
  freeze: auto
tbl-cap-location: bottom
fig-cap-location: bottom
---

::: {.callout-important}
## Scope and assumptions

This tutorial is configured for a **binary bulk phenotype**, with the **Neural-biased Morphotype encoded as `1`** and all other morphotypes encoded as `0`. Thus, `Positive` cells are interpreted as associated with the Neural-biased phenotype, while `Negative` cells are associated with the non-neural reference group.

The single-cell object already contains biological cell-type/state labels in
`luque_cluster_annotation`; no cell-type annotation step is run. Mouse gene
symbols are already present, so gene-symbol conversion is disabled.

The four screening methods are **scPP**, **LP-SGL**, **DEGAS**, and
**Scissor** only.
:::

# 0. Configuration: files, columns, and all hyperparameters

All paths, column names, randomisation controls, thresholds, and
method-specific settings are defined here. Change this section rather than
editing values later in the workflow.

## 0.1 File locations

```{r}
#| label: configuration-paths

PATHS <- list(
  # Input files ---------------------------------------------------------------
  sc_seurat = file.path("data", "single_cell", "luque_seurat.rds"),
  bulk_summarized_experiment = file.path(
    "data", "bulk", "bulk_counts_summarized_experiment.rds"
  ),

  # Required only because Scissor cell evaluation is enabled.
  # This must be an .RData benchmark file accepted by DoScissorCellEval().
  scissor_benchmark = file.path(
    "data", "benchmarks", "Scissor_benchmark.RData"
  ),

  # Persistent project outputs -----------------------------------------------
  output_dir = file.path("results", "SigBridgeR_neural_morphotype"),
  figures_dir = file.path(
    "results", "SigBridgeR_neural_morphotype", "figures"
  ),
  tables_dir = file.path(
    "results", "SigBridgeR_neural_morphotype", "tables"
  ),
  objects_dir = file.path(
    "results", "SigBridgeR_neural_morphotype", "objects"
  ),

  # Temporary/cached intermediates. These disappear with the R session.
  tmp_root = file.path(tempdir(), "SigBridgeR_neural_morphotype"),

  # Local uv project for DEGAS Python dependencies ----------------------------
  degas_uv_project = file.path(".uv", "degas")
)

PATHS$degas_venv <- file.path(PATHS$degas_uv_project, ".venv")
PATHS$degas_python <- if (.Platform$OS.type == "windows") {
  file.path(PATHS$degas_venv, "Scripts", "python.exe")
} else {
  file.path(PATHS$degas_venv, "bin", "python")
}

PATHS$preprocessed_sc <- file.path(
  PATHS$objects_dir, "single_cell_preprocessed.rds"
)
PATHS$preprocessed_bulk_counts <- file.path(
  PATHS$objects_dir, "bulk_filtered_counts.rds"
)
PATHS$preprocessed_bulk_expression <- file.path(
  PATHS$objects_dir, "bulk_logCPM_for_screening.rds"
)
PATHS$phenotype <- file.path(
  PATHS$objects_dir, "neural_biased_phenotype.rds"
)
PATHS$screen_results <- file.path(
  PATHS$objects_dir, "screen_results.rds"
)
PATHS$merged_result <- file.path(
  PATHS$objects_dir, "merged_weighted_vote.rds"
)

invisible(lapply(
  unique(c(
    PATHS$output_dir,
    PATHS$figures_dir,
    PATHS$tables_dir,
    PATHS$objects_dir,
    PATHS$tmp_root,
    PATHS$degas_uv_project
  )),
  dir.create,
  recursive = TRUE,
  showWarnings = FALSE
))
```

## 0.2 Data columns and global controls

```{r}
#| label: configuration-columns-global

COL <- list(
  # Existing single-cell metadata annotation
  cell_annotation = "luque_cluster_annotation",

  # SummarizedExperiment::colData() column containing the bulk morphotype
  bulk_morphotype = "Morphotype",

  # Optional batch column in colData(); use NULL when absent
  bulk_batch = NULL,

  # Assay containing raw bulk counts
  bulk_assay = "counts",

  # Input and output Seurat assays
  sc_input_assay = "RNA",
  sc_analysis_assay = "SCT"
)

CFG <- list(
  project_name = "SigBridgeR_NeuralBiased",
  phenotype_name = "NeuralBiasedMorphotype",
  positive_label = "Neural-biased Morphotype",
  positive_code = 1L,
  reference_code = 0L,
  phenotype_class = "binary",

  seed = 20260713L,
  cv_folds = 5L,
  bootstrap_n = 100L,

  p_value = 0.05,
  fdr = 0.05,
  log2fc = 0.5,
  penalty_mix = 0.5,
  selection_fraction = 0.20,

  min_common_genes = 500L,
  minimum_samples_per_class = 5L,

  # Half of all currently available cores
  cpu_budget = max(
    1L,
    floor(parallelly::availableCores() / 2L)
  ),
  future_globals_max_bytes = 16 * 1024^3,

  # DEGAS: uv/pip is more likely to resolve TensorFlow 2.4.1 with Python 3.8.
  # Override this path/version for platforms without a compatible wheel.
  degas_python_version = "3.8.18",
  degas_python_packages = c(
    "tensorflow==2.4.1",
    "protobuf==3.20.3",
    "numpy==1.19.5"
  )
)

CFG$n_r_workers <- min(3L, CFG$cpu_budget)
CFG$threads_per_worker <- max(
  1L,
  floor(CFG$cpu_budget / CFG$n_r_workers)
)

set.seed(CFG$seed)
```

## 0.3 Preprocessing hyperparameters

There is **no explicit `ScaleData()` step**. `SCTransform()` performs its own
variance-stabilising transformation and stores Pearson residuals in the SCT
assay, but the separate Seurat scaling step is omitted.

Mitochondrial and ribosomal percentages are calculated for QC. They are not
used as automatic exclusion thresholds because mitochondrial and ribosomal
programmes can be biologically meaningful in developmental cell lines.

```{r}
#| label: configuration-preprocessing

PREPROCESS <- list(
  single_cell = list(
    # Create -> SCTransform -> PCA -> neighbours -> UMAP.
    # No ScaleData and no re-clustering.
    pipeline = "orpeu",

    create = list(
      project = CFG$project_name,
      min.cells = 400L,
      min.features = 0L
    ),

    sctransform = list(
      vst.flavor = "v2",
      variable.features.n = 3000L,
      return.only.var.genes = FALSE,
      vars.to.regress = NULL,
      conserve.memory = FALSE,
      verbose = FALSE
    ),

    pca = list(
      npcs = 50L,
      verbose = FALSE
    ),

    neighbours = list(
      dims = 1:30,
      verbose = FALSE
    ),

    umap = list(
      dims = 1:30,
      seed.use = CFG$seed,
      verbose = FALSE
    ),

    # Mouse-style symbols
    qc_patterns = c("^mt-", "^Rp[sl]"),

    # QCFilter merges its own defaults into this list. Explicit thresholds
    # above 100 disable mitochondrial/ribosomal exclusion while retaining the
    # calculated QC columns and their reports.
    data_filter = list(
      assay = "RNA",
      nFeature_thresh = c(200L, 6000L),
      nCount_thresh = c(500L, 50000L),
      percent.mt = 101,
      percent.rp = 101
    )
  ),

  bulk = list(
    gene_symbol_conversion = FALSE,
    check = TRUE,
    min_count_threshold = 10L,
    min_gene_expressed = 3L,
    min_total_reads = 1e6L,
    min_genes_detected = 10000L,

    # Keep the correlation QC report, but do not remove samples by correlation.
    # Pearson correlations cannot be lower than -1.
    min_correlation = -1,

    n_top_genes = 500L,
    show_plot_results = TRUE,

    # Screening expression transform; this is normalization, not gene-ID
    # standardization or per-gene z-scoring.
    screening_transform = "edgeR_logCPM",
    log_cpm_prior_count = 1
  )
)
```

::: {.callout-note}
`min.cells = 400` is retained because it is the SigBridgeR default. For a
small dataset or a study focused on rare developmental states, review the
number of genes removed and lower this value only through the configuration
section; do not change it silently downstream.
:::

## 0.4 Screening hyperparameters

```{r}
#| label: configuration-screening

SCREEN <- list(
  shared = list(
    assay = COL$sc_analysis_assay,
    seed = CFG$seed,
    verbose = TRUE,
    cv_folds = CFG$cv_folds,
    bootstrap_n = CFG$bootstrap_n,
    p_value = CFG$p_value,
    fdr = CFG$fdr,
    log2fc = CFG$log2fc,
    penalty_mix = CFG$penalty_mix,
    selection_fraction = CFG$selection_fraction
  ),

  scPP = list(
    ref_group = CFG$reference_code,
    Log2FC_cutoff = CFG$log2fc,
    estimate_cutoff = 0.2,
    probs = CFG$selection_fraction,
    parallel = TRUE
  ),

  LP_SGL = list(
    family = "logit",
    resolution = 0.6,
    alpha = CFG$penalty_mix,
    nfold = CFG$cv_folds,
    dge_analysis = list(
      run = TRUE,
      logFC_threshold = CFG$log2fc,
      pval_threshold = CFG$p_value
    )
  ),

  Scissor = list(
    family = "binomial",

    # Equal emphasis on L1 and network penalties.
    alpha = CFG$penalty_mix,

    # Retained for completeness. In Scissor, cutoff is used only when
    # alpha = NULL, so it is not active with alpha = 0.5.
    cutoff = CFG$selection_fraction,

    reliability_test = list(
      run = TRUE,
      n = CFG$bootstrap_n,
      nfold = CFG$cv_folds
    ),

    cell_evaluation = list(
      run = TRUE,
      benchmark_data = PATHS$scissor_benchmark,
      FDR_cutoff = CFG$fdr,
      bootstrap_n = CFG$bootstrap_n
    ),

    intermediate_dir = PATHS$tmp_root
  ),

  DEGAS = list(
    phenotype_class = "binary",
    sc_data.pheno_colname = COL$cell_annotation,

    select_fraction = CFG$selection_fraction,
    min_thresh = 0.4,
    normality_test_method = "jarque-bera",

    # Correct interpretation of the requested "StandardClass + FeedForward":
    # ClassClass uses both the cell annotations and bulk binary labels.
    # Standard is the feed-forward architecture.
    degas_params = list(
      DEGAS.model_type = "ClassClass",
      DEGAS.architecture = "Standard",
      DEGAS.ff_depth = 3L,
      DEGAS.bag_depth = CFG$bootstrap_n,
      DEGAS.pyloc = PATHS$degas_python,
      DEGAS.train_steps = 2000L,
      DEGAS.scbatch_sz = 200L,
      DEGAS.patbatch_sz = 50L,
      DEGAS.hidden_feats = 50L,
      DEGAS.do_prc = 0.5,

      # Equal coefficients, as requested. These are regularisation
      # coefficients, not probabilities that must sum to one.
      DEGAS.lambda1 = CFG$penalty_mix,
      DEGAS.lambda2 = CFG$penalty_mix,
      DEGAS.lambda3 = CFG$penalty_mix,

      DEGAS.seed = CFG$seed
    )
  ),

  consensus = list(
    method_weights = c(
      scPP = 0.25,
      LP_SGL = 0.25,
      DEGAS = 0.25,
      Scissor = 0.25
    ),
    ties.method = "random"
  )
)
```

::: {.callout-warning}
## Limits of cross-method harmonisation

The four methods do not expose equivalent controls.

- A literal common fold assignment cannot be supplied: LP-SGL and the Scissor
  reliability procedure accept only `nfold`, not a shared `foldid`. The same
  seed and five folds are therefore used wherever supported.
- scPP and DEGAS do not expose five-fold CV through the SigBridgeR interface.
- The 20% threshold maps to `scPP::probs` and
  `DEGAS::select_fraction`. Scissor's `cutoff` is inactive when a fixed
  `alpha` is supplied, and LP-SGL has no corresponding cell-fraction cutoff.
- FDR, p-value, and log-fold-change thresholds are passed only to algorithms
  that expose them.
:::

## 0.5 Display the complete configuration

```{r}
#| label: configuration-table

flatten_config <- function(x, parent = "") {
  purrr::imap_dfr(x, function(value, name) {
    full_name <- if (nzchar(parent)) paste(parent, name, sep = ".") else name

    if (is.list(value)) {
      return(flatten_config(value, full_name))
    }

    tibble::tibble(
      parameter = full_name,
      value = if (is.null(value)) {
        "NULL"
      } else {
        paste(as.character(value), collapse = ", ")
      }
    )
  })
}

config_table <- dplyr::bind_rows(
  flatten_config(PATHS, "PATHS"),
  flatten_config(COL, "COL"),
  flatten_config(CFG, "CFG"),
  flatten_config(PREPROCESS, "PREPROCESS"),
  flatten_config(SCREEN, "SCREEN")
)

knitr::kable(config_table, caption = "Complete pipeline configuration")
```

## 0.6 One-time package installation

Run this chunk only when creating the project environment. The four
algorithm packages are the remotes declared by the current SigBridgeR
package.

```{r}
#| label: package-installation
#| eval: false

install.packages(c(
  "pak",
  "BiocManager",
  "future",
  "future.apply",
  "parallelly",
  "ggplot2",
  "patchwork",
  "ggupset",
  "readr",
  "knitr"
))

BiocManager::install(c(
  "SummarizedExperiment",
  "edgeR",
  "limma"
))

pak::pkg_install(c(
  "WangLabCSU/SigBridgeR",
  "Exceret/DEGAS",
  "Exceret/LPSGL",
  "Exceret/Scissor",
  "Exceret/ScPP"
))
```

## 0.7 Packages and reproducible execution settings

```{r}
#| label: packages

required_packages <- c(
  "SigBridgeR",
  "SigBridgeRUtils",
  "Seurat",
  "SeuratObject",
  "SummarizedExperiment",
  "S4Vectors",
  "Matrix",
  "data.table",
  "dplyr",
  "tibble",
  "purrr",
  "ggplot2",
  "patchwork",
  "future",
  "future.apply",
  "parallelly",
  "reticulate",
  "DEGAS",
  "ScPP",
  "LPSGL",
  "Scissor",
  "limma",
  "edgeR",
  "ggupset",
  "readr",
  "knitr"
)

missing_packages <- required_packages[
  !vapply(required_packages, requireNamespace, logical(1), quietly = TRUE)
]

if (length(missing_packages) > 0L) {
  stop(
    "Install the following packages before running the pipeline: ",
    paste(missing_packages, collapse = ", ")
  )
}

suppressPackageStartupMessages({
  library(SigBridgeR)
  library(Seurat)
  library(SummarizedExperiment)
  library(Matrix)
  library(data.table)
  library(dplyr)
  library(purrr)
  library(ggplot2)
  library(patchwork)
})

if (utils::packageVersion("SigBridgeR") < "3.8.1") {
  warning(
    "This tutorial targets the current 3.8.1-era API. ",
    "Update SigBridgeR or re-check function signatures."
  )
}

options(future.globals.maxSize = CFG$future_globals_max_bytes)
set.seed(CFG$seed)

# Configure no more than half the available cores before TensorFlow is imported.
SigBridgeR::setThreads(
  threads = CFG$cpu_budget,
  tf_config = list(
    inter_op = max(1L, floor(CFG$cpu_budget / 4L)),
    intra_op = CFG$cpu_budget
  ),
  verbose = TRUE
)

package_versions <- tibble::tibble(
  package = required_packages,
  version = vapply(
    required_packages,
    function(pkg) as.character(utils::packageVersion(pkg)),
    character(1)
  )
)

knitr::kable(package_versions, caption = "Package versions")
```

# 1. Data preprocessing

::: {.callout-note appearance="simple"}
## Preprocessing design

1. Map the Neural-biased bulk morphotype to `1`.
2. Rebuild the single-cell analysis from the raw-count layer of the original
   Seurat object while retaining all original metadata.
3. Use `SCTransform()` without a separate `ScaleData()` step.
4. Calculate mouse mitochondrial and ribosomal QC metrics, but do not exclude
   cells using those percentages.
5. Process raw bulk counts from a `SummarizedExperiment`.
6. Run `PhenoPreProcess()` last to validate and align bulk samples and labels.
:::

## 1.1 Load and validate the two source objects

```{r}
#| label: load-inputs

stopifnot(file.exists(PATHS$sc_seurat))
stopifnot(file.exists(PATHS$bulk_summarized_experiment))

sc_original <- readRDS(PATHS$sc_seurat)
bulk_se <- readRDS(PATHS$bulk_summarized_experiment)

stopifnot(inherits(sc_original, "Seurat"))
stopifnot(inherits(bulk_se, "SummarizedExperiment"))
stopifnot(COL$cell_annotation %in% colnames(sc_original[[]]))
stopifnot(COL$bulk_morphotype %in% colnames(SummarizedExperiment::colData(bulk_se)))
stopifnot(COL$bulk_assay %in% SummarizedExperiment::assayNames(bulk_se))

if (anyNA(sc_original[[]][[COL$cell_annotation]])) {
  stop("Missing values detected in `", COL$cell_annotation, "`.")
}
```

## 1.2 Build the binary phenotype with `PhenoMap()`

```{r}
#| label: phenotype-map

bulk_counts_raw <- SummarizedExperiment::assay(
  bulk_se,
  COL$bulk_assay
)
bulk_counts_raw <- as.matrix(bulk_counts_raw)

if (is.null(colnames(bulk_counts_raw)) || anyDuplicated(colnames(bulk_counts_raw))) {
  stop("Bulk sample names must be present and unique.")
}

bulk_coldata <- as.data.frame(
  SummarizedExperiment::colData(bulk_se)
)

morphotype_raw <- stats::setNames(
  as.character(bulk_coldata[[COL$bulk_morphotype]]),
  colnames(bulk_counts_raw)
)

phenotype_mapped <- SigBridgeR::PhenoMap(
  morphotype_raw,
  morphotype_raw == CFG$positive_label ~ CFG$positive_code,
  .default = CFG$reference_code
)

phenotype_mapped <- as.integer(phenotype_mapped)
names(phenotype_mapped) <- colnames(bulk_counts_raw)

phenotype_counts <- table(
  Morphotype = morphotype_raw,
  Encoded = phenotype_mapped
)

phenotype_counts
```

```{r}
#| label: phenotype-balance

class_counts <- table(phenotype_mapped)

if (length(class_counts) != 2L) {
  stop("The mapped bulk phenotype is not binary.")
}

if (any(class_counts < CFG$minimum_samples_per_class)) {
  stop(
    "At least one phenotype class contains fewer than ",
    CFG$minimum_samples_per_class,
    " samples."
  )
}

class_fractions <- prop.table(class_counts)
class_fractions

if (diff(range(class_fractions)) > 0.10) {
  warning(
    "The bulk classes differ by more than 10 percentage points; ",
    "reconsider whether class weighting is required."
  )
}
```

## 1.3 Single-cell preprocessing with `SCPreProcess()`

Passing an existing Seurat object directly to `SCPreProcess()` only validates
and optionally repairs it. To rerun the requested SigBridgeR preprocessing,
the raw-count layer and the existing metadata are passed explicitly.

```{r}
#| label: single-cell-preprocess

sc_counts <- SeuratObject::LayerData(
  sc_original,
  assay = COL$sc_input_assay,
  layer = "counts"
)

sc_metadata <- sc_original[[]]
sc_metadata <- sc_metadata[colnames(sc_counts), , drop = FALSE]

sc_data <- SigBridgeR::SCPreProcess(
  sc = sc_counts,
  pipeline = PREPROCESS$single_cell$pipeline,

  params = list(
    o = c(
      PREPROCESS$single_cell$create,
      list(meta.data = sc_metadata)
    ),
    r = PREPROCESS$single_cell$sctransform,
    p = PREPROCESS$single_cell$pca,
    e = PREPROCESS$single_cell$neighbours,
    u = PREPROCESS$single_cell$umap
  ),

  quality_control = list(
    pattern = PREPROCESS$single_cell$qc_patterns
  ),

  data_filter = PREPROCESS$single_cell$data_filter,
  column2only_tumor = NULL,
  verbose = TRUE
)

SeuratObject::DefaultAssay(sc_data) <- COL$sc_analysis_assay

stopifnot(COL$cell_annotation %in% colnames(sc_data[[]]))
stopifnot("umap" %in% names(sc_data@reductions))

saveRDS(sc_data, PATHS$preprocessed_sc)
```

### Single-cell QC report

```{r}
#| label: single-cell-qc
#| fig-width: 12
#| fig-height: 8

qc_columns <- intersect(
  c(
    "nFeature_RNA",
    "nCount_RNA",
    "percent.mt",
    "percent.rp",
    "percent.mt_rp"
  ),
  colnames(sc_data[[]])
)

if (length(qc_columns) > 0L) {
  print(
    Seurat::VlnPlot(
      sc_data,
      features = qc_columns,
      group.by = COL$cell_annotation,
      pt.size = 0,
      ncol = min(3L, length(qc_columns))
    ) +
      patchwork::plot_annotation(
        title = "Single-cell QC by existing biological annotation"
      )
  )
}

sc_qc_summary <- sc_data[[]] |>
  dplyr::group_by(.data[[COL$cell_annotation]]) |>
  dplyr::summarise(
    n_cells = dplyr::n(),
    dplyr::across(
      dplyr::all_of(qc_columns),
      list(
        median = ~ stats::median(.x, na.rm = TRUE),
        q05 = ~ stats::quantile(.x, 0.05, na.rm = TRUE),
        q95 = ~ stats::quantile(.x, 0.95, na.rm = TRUE)
      ),
      .names = "{.col}_{.fn}"
    ),
    .groups = "drop"
  )

readr::write_csv(
  sc_qc_summary,
  file.path(PATHS$tables_dir, "single_cell_qc_summary.csv")
)

knitr::kable(sc_qc_summary, caption = "Single-cell QC summary")
```

## 1.4 Bulk preprocessing from raw `SummarizedExperiment` counts

`min_correlation = -1` leaves the correlation QC calculation and PCA report
enabled while preventing correlation-based sample removal. `BulkPreProcess()`
returns filtered counts, which are preserved as a separate object. A log-CPM
matrix is then created for phenotype screening; no per-gene z-scoring is
applied.

```{r}
#| label: bulk-preprocess
#| fig-width: 8
#| fig-height: 6

sample_info <- data.frame(
  sample = colnames(bulk_counts_raw),
  condition = phenotype_mapped,
  stringsAsFactors = FALSE
)

if (
  !is.null(COL$bulk_batch) &&
  COL$bulk_batch %in% colnames(bulk_coldata)
) {
  sample_info$batch <- bulk_coldata[
    match(sample_info$sample, rownames(bulk_coldata)),
    COL$bulk_batch
  ]
}

bulk_counts_filtered <- SigBridgeR::BulkPreProcess(
  data = list(
    count_matrix = bulk_counts_raw,
    sample_info = sample_info
  ),
  gene_symbol_conversion = PREPROCESS$bulk$gene_symbol_conversion,
  check = PREPROCESS$bulk$check,
  min_count_threshold = PREPROCESS$bulk$min_count_threshold,
  min_gene_expressed = PREPROCESS$bulk$min_gene_expressed,
  min_total_reads = PREPROCESS$bulk$min_total_reads,
  min_genes_detected = PREPROCESS$bulk$min_genes_detected,
  min_correlation = PREPROCESS$bulk$min_correlation,
  n_top_genes = PREPROCESS$bulk$n_top_genes,
  show_plot_results = PREPROCESS$bulk$show_plot_results,
  seed = CFG$seed,
  verbose = TRUE
)

# Keep the filtered integer counts for auditability and count-aware follow-up.
saveRDS(
  bulk_counts_filtered,
  PATHS$preprocessed_bulk_counts
)

# SigBridgeR's screening examples use transformed expression. Convert the
# filtered counts to log2 counts-per-million without per-gene z-scoring.
matched_bulk <- edgeR::cpm(
  bulk_counts_filtered,
  log = TRUE,
  prior.count = PREPROCESS$bulk$log_cpm_prior_count,
  normalized.lib.sizes = TRUE
)

saveRDS(
  matched_bulk,
  PATHS$preprocessed_bulk_expression
)
```

### Bulk QC summary

```{r}
#| label: bulk-qc

bulk_qc_summary <- tibble::tibble(
  sample = colnames(bulk_counts_filtered),
  total_reads = Matrix::colSums(bulk_counts_filtered),
  genes_detected = Matrix::colSums(bulk_counts_filtered > 0),
  phenotype = phenotype_mapped[colnames(bulk_counts_filtered)]
)

readr::write_csv(
  bulk_qc_summary,
  file.path(PATHS$tables_dir, "bulk_qc_summary.csv")
)

knitr::kable(bulk_qc_summary, caption = "Bulk QC summary after preprocessing")
```

## 1.5 Final alignment and validation with `PhenoPreProcess()`

```{r}
#| label: phenotype-preprocess-validation

phenotype <- SigBridgeR::PhenoPreProcess(
  bulk = matched_bulk,
  phenotype = phenotype_mapped,
  phenotype_class = CFG$phenotype_class,
  verbose = TRUE
)

# PhenoPreProcess retains common samples. Reorder both bulk representations.
matched_bulk <- matched_bulk[, names(phenotype), drop = FALSE]
bulk_counts_filtered <- bulk_counts_filtered[
  ,
  names(phenotype),
  drop = FALSE
]

stopifnot(
  identical(colnames(matched_bulk), names(phenotype)),
  length(unique(phenotype)) == 2L,
  !anyNA(matched_bulk),
  !anyNA(phenotype)
)

sc_expression <- SeuratObject::LayerData(
  sc_data,
  assay = COL$sc_analysis_assay,
  layer = "data"
)

bulk_genes_after_filtering <- nrow(matched_bulk)

common_genes <- intersect(
  rownames(matched_bulk),
  rownames(sc_expression)
)

if (length(common_genes) < CFG$min_common_genes) {
  stop(
    "Only ", length(common_genes),
    " genes overlap between bulk and single-cell data; expected at least ",
    CFG$min_common_genes, "."
  )
}

matched_bulk <- matched_bulk[common_genes, , drop = FALSE]
bulk_counts_filtered <- bulk_counts_filtered[
  intersect(common_genes, rownames(bulk_counts_filtered)),
  ,
  drop = FALSE
]

saveRDS(
  bulk_counts_filtered,
  PATHS$preprocessed_bulk_counts
)
saveRDS(
  matched_bulk,
  PATHS$preprocessed_bulk_expression
)
saveRDS(phenotype, PATHS$phenotype)

input_validation <- tibble::tibble(
  item = c(
    "Single cells",
    "Bulk samples",
    "Bulk genes after filtering",
    "Shared genes used",
    "Positive bulk samples",
    "Reference bulk samples"
  ),
  value = c(
    ncol(sc_data),
    ncol(matched_bulk),
    bulk_genes_after_filtering,
    length(common_genes),
    sum(phenotype == CFG$positive_code),
    sum(phenotype == CFG$reference_code)
  )
)

knitr::kable(input_validation, caption = "Validated SigBridgeR inputs")
```

# 2. Screen with scPP, LP-SGL, DEGAS, and Scissor

::: {.callout-important}
## DEGAS model choice

`StandardClass` is not a valid DEGAS model type. Because both the existing
single-cell labels and the binary bulk labels are used, this workflow sets:

- `DEGAS.model_type = "ClassClass"`
- `DEGAS.architecture = "Standard"`

The latter is DEGAS's standard feed-forward network rather than DenseNet.
:::

## 2.1 Create or reuse the local DEGAS `uv` environment

SigBridgeR's environment helper officially recognises Conda and `venv`, not
`uv`. The workflow therefore lets `uv` create the local virtual environment
and passes the resulting Python executable directly to DEGAS.

```{r}
#| label: degas-uv-environment

run_system_checked <- function(command, args) {
  output <- system2(
    command,
    args = args,
    stdout = TRUE,
    stderr = TRUE
  )

  status <- attr(output, "status")
  if (is.null(status)) {
    status <- 0L
  }

  if (!identical(as.integer(status), 0L)) {
    stop(
      "Command failed: ",
      command, " ", paste(args, collapse = " "),
      "\n", paste(output, collapse = "\n")
    )
  }

  invisible(output)
}

if (!file.exists(PATHS$degas_python)) {
  uv_path <- Sys.which("uv")

  if (!nzchar(uv_path)) {
    stop(
      "`uv` is not installed or is not on PATH. ",
      "Install uv, or set PATHS$degas_python to an existing compatible Python."
    )
  }

  dir.create(
    PATHS$degas_uv_project,
    recursive = TRUE,
    showWarnings = FALSE
  )

  run_system_checked(
    uv_path,
    c(
      "venv",
      "--python", CFG$degas_python_version,
      PATHS$degas_venv
    )
  )

  run_system_checked(
    uv_path,
    c(
      "pip", "install",
      "--python", PATHS$degas_python,
      CFG$degas_python_packages
    )
  )
}

stopifnot(file.exists(PATHS$degas_python))

reticulate::use_python(PATHS$degas_python, required = TRUE)
reticulate::py_config()
```

::: {.callout-caution}
TensorFlow 2.4.1 is an old dependency and compatible wheels are
platform-specific. On platforms where `uv` cannot resolve it, point
`PATHS$degas_python` to a tested local Python environment and keep the rest of
the workflow unchanged.
:::

## 2.2 Define validated screening executors

The wrappers expose the common SigBridgeR screening interface, keep all
method-specific settings in the first section, and return a literal list with
`scRNA_data` as required by `ValidateScreenFunc()`.

### scPP executor

```{r}
#| label: executor-scpp

screen_scPP_core <- function(
    sc_data,
    matched_bulk,
    phenotype,
    label_type,
    phenotype_class,
    verbose = TRUE,
    ...) {

  out <- SigBridgeR::DoscPP(
    matched_bulk = matched_bulk,
    sc_data = sc_data,
    phenotype = phenotype,
    label_type = label_type,
    phenotype_class = phenotype_class,
    ref_group = SCREEN$scPP$ref_group,
    Log2FC_cutoff = SCREEN$scPP$Log2FC_cutoff,
    estimate_cutoff = SCREEN$scPP$estimate_cutoff,
    probs = SCREEN$scPP$probs,
    assay = SCREEN$shared$assay,
    seed = SCREEN$shared$seed,
    parallel = SCREEN$scPP$parallel,
    verbose = verbose
  )

  list(
    scRNA_data = out$scRNA_data,
    result = out
  )
}
```

### LP-SGL executor with differential-expression analysis

```{r}
#| label: executor-lpsgl

screen_LP_SGL_core <- function(
    sc_data,
    matched_bulk,
    phenotype,
    label_type,
    phenotype_class,
    verbose = TRUE,
    ...) {

  if (!identical(phenotype_class, "binary")) {
    stop("This configured LP-SGL executor expects a binary phenotype.")
  }

  out <- SigBridgeR::DoLP_SGL(
    matched_bulk = matched_bulk,
    sc_data = sc_data,
    phenotype = phenotype,
    label_type = label_type,
    family = SCREEN$LP_SGL$family,
    resolution = SCREEN$LP_SGL$resolution,
    alpha = SCREEN$LP_SGL$alpha,
    nfold = SCREEN$LP_SGL$nfold,
    dge_analysis = SCREEN$LP_SGL$dge_analysis,
    assay = SCREEN$shared$assay,
    seed = SCREEN$shared$seed,
    verbose = verbose
  )

  list(
    scRNA_data = out$scRNA_data,
    result = out
  )
}
```

### Scissor executor with reliability testing and cell evaluation

This wrapper calls the same Scissor backend and QC functions used by
SigBridgeR, but writes the first-run intermediate `.RData` file directly under
`tempdir()`. This avoids the deprecated cache argument names shown in older
vignettes and keeps the tutorial independent of cache-path fallback behaviour.

```{r}
#| label: executor-scissor

screen_Scissor_core <- function(
    sc_data,
    matched_bulk,
    phenotype,
    label_type,
    phenotype_class,
    verbose = TRUE,
    ...) {

  if (!identical(phenotype_class, "binary")) {
    stop("This configured Scissor executor expects a binary phenotype.")
  }

  if (
    isTRUE(SCREEN$Scissor$cell_evaluation$run) &&
    !file.exists(SCREEN$Scissor$cell_evaluation$benchmark_data)
  ) {
    stop(
      "Scissor cell evaluation is enabled, but the benchmark file does not exist: ",
      SCREEN$Scissor$cell_evaluation$benchmark_data
    )
  }

  dir.create(
    SCREEN$Scissor$intermediate_dir,
    recursive = TRUE,
    showWarnings = FALSE
  )

  scissor_input_file <- tempfile(
    pattern = "Scissor_inputs_",
    tmpdir = SCREEN$Scissor$intermediate_dir,
    fileext = ".RData"
  )

  tag <- c(
    paste0(label_type, "_Negative"),
    paste0(label_type, "_Positive")
  )

  set.seed(SCREEN$shared$seed)

  infos <- Scissor::Scissor.v5.optimized(
    bulk_dataset = matched_bulk,
    sc_dataset = sc_data,
    phenotype = phenotype,
    tag = tag,
    alpha = SCREEN$Scissor$alpha,
    cutoff = SCREEN$Scissor$cutoff,
    family = SCREEN$Scissor$family,
    Save_file = scissor_input_file,
    Load_file = NULL,
    assay = SCREEN$shared$assay,
    seed = SCREEN$shared$seed,
    verbose = verbose
  )

  scissor_metadata <- data.frame(
    scissor = rep("Neutral", ncol(sc_data)),
    row.names = colnames(sc_data)
  )
  scissor_metadata$scissor[
    rownames(scissor_metadata) %in% infos$Scissor_pos
  ] <- "Positive"
  scissor_metadata$scissor[
    rownames(scissor_metadata) %in% infos$Scissor_neg
  ] <- "Negative"

  selected_cell_n <- length(infos$Scissor_pos) + length(infos$Scissor_neg)

  set.seed(SCREEN$shared$seed)

  reliability_result <- if (
    isTRUE(SCREEN$Scissor$reliability_test$run) &&
    length(unique(infos$Y)) >= 2L &&
    selected_cell_n > 0L
  ) {
    Scissor::reliability.test(
      X = infos$X,
      Y = infos$Y,
      network = infos$network,
      alpha = infos$para$alpha,
      family = SCREEN$Scissor$family,
      cell_num = selected_cell_n,
      n = SCREEN$Scissor$reliability_test$n,
      nfold = SCREEN$Scissor$reliability_test$nfold,
      verbose = verbose
    )
  } else {
    NULL
  }

  set.seed(SCREEN$shared$seed)

  cell_evaluation <- if (
    isTRUE(SCREEN$Scissor$cell_evaluation$run)
  ) {
    Scissor::evaluate.cell(
      Load_file = SCREEN$Scissor$cell_evaluation$benchmark_data,
      Scissor_result = infos,
      FDR_cutoff = SCREEN$Scissor$cell_evaluation$FDR_cutoff,
      bootstrap_n = SCREEN$Scissor$cell_evaluation$bootstrap_n
    )
  } else {
    NULL
  }

  sc_out <- SeuratObject::AddMetaData(
    sc_data,
    metadata = scissor_metadata
  )

  sc_out <- SigBridgeRUtils::AddMisc(
    seurat_obj = sc_out,
    scissor_type = label_type,
    scissor_para = c(
      infos$para,
      list(reliability_test = reliability_result)
    ),
    cover = FALSE
  )

  list(
    scRNA_data = sc_out,
    result = list(
      scRNA_data = sc_out,
      scissor_result = infos,
      reliability_result = reliability_result,
      cell_evaluation = cell_evaluation,
      intermediate_file = scissor_input_file
    )
  )
}
```

### uv-backed DEGAS executor

The current public `DoDEGAS()` interface still contains a soft-deprecated
`tmp_dir` argument and couples it to cache fallback logic. The following thin
adapter calls the same current DEGAS backend used by SigBridgeR while:

1. using the local `uv` Python executable;
2. keeping transient model files under `tempdir()`;
3. using `ClassClass` and the `Standard` feed-forward architecture;
4. returning the same main output structure.

```{r}
#| label: executor-degas-backend

run_DEGAS_uv_backend <- function(
    sc_data,
    matched_bulk,
    phenotype,
    label_type,
    phenotype_class,
    verbose = TRUE) {

  if (!identical(phenotype_class, "binary")) {
    stop("This configured DEGAS executor expects a binary phenotype.")
  }

  p <- SCREEN$DEGAS
  d <- p$degas_params

  stopifnot(file.exists(d$DEGAS.pyloc))
  reticulate::use_python(d$DEGAS.pyloc, required = TRUE)

  sc_meta <- sc_data[[]]
  if (!p$sc_data.pheno_colname %in% colnames(sc_meta)) {
    stop(
      "DEGAS single-cell label column is missing: ",
      p$sc_data.pheno_colname
    )
  }

  pheno_df <- DEGAS::Vec2sparse(
    phenotype,
    col_prefix = label_type
  )

  sc_pheno <- DEGAS::Vec2sparse(
    sc_meta[[p$sc_data.pheno_colname]]
  )

  sc_mat <- SeuratObject::LayerData(
    sc_data,
    assay = SCREEN$shared$assay,
    layer = "data"
  )

  common <- intersect(rownames(matched_bulk), rownames(sc_mat))
  if (length(common) < CFG$min_common_genes) {
    stop("Too few shared genes for DEGAS: ", length(common))
  }

  t_sc_mat <- Matrix::t(sc_mat[common, , drop = FALSE])
  t_bulk <- Matrix::t(matched_bulk[common, , drop = FALSE])

  degas_tmp <- tempfile(
    pattern = "DEGAS_",
    tmpdir = PATHS$tmp_root
  )
  dir.create(degas_tmp, recursive = TRUE, showWarnings = FALSE)
  degas_tmp <- paste0(normalizePath(degas_tmp), .Platform$file.sep)

  # DEGAS's current backend reads several settings from the global environment.
  d$path.data <- degas_tmp
  d$path.result <- degas_tmp
  degas_package_root <- system.file(package = "DEGAS")
  d$DEGAS.toolsPath <- paste0(
    file.path(degas_package_root, "DEGAS_tools"),
    .Platform$file.sep
  )

  global_names <- names(d)
  existed_before <- vapply(
    global_names,
    exists,
    logical(1),
    envir = .GlobalEnv,
    inherits = FALSE
  )

  old_globals <- if (any(existed_before)) {
    mget(
      global_names[existed_before],
      envir = .GlobalEnv,
      inherits = FALSE
    )
  } else {
    list()
  }

  list2env(d, envir = .GlobalEnv)

  on.exit({
    current_globals <- intersect(global_names, ls(envir = .GlobalEnv))
    if (length(current_globals) > 0L) {
      rm(list = current_globals, envir = .GlobalEnv)
    }
    if (length(old_globals) > 0L) {
      list2env(old_globals, envir = .GlobalEnv)
    }
    unlink(degas_tmp, recursive = TRUE, force = TRUE)
    gc(verbose = FALSE)
  }, add = TRUE)

  set.seed(SCREEN$shared$seed)

  if (verbose) {
    message(
      "Training DEGAS: ClassClass + Standard, ",
      d$DEGAS.bag_depth, " bootstrap models."
    )
  }

  model <- DEGAS::runCCMTLBag.optimized(
    scExp = t_sc_mat,
    scLab = sc_pheno,
    patExp = t_bulk,
    patLab = pheno_df,
    tmpDir = degas_tmp,
    model_type = d$DEGAS.model_type,
    architecture = d$DEGAS.architecture,
    FFdepth = d$DEGAS.ff_depth,
    Bagdepth = d$DEGAS.bag_depth,
    DEGAS.seed = d$DEGAS.seed,
    verbose = verbose
  )

  names(model) <- paste0(
    "ccModel_",
    seq_len(d$DEGAS.bag_depth)
  )

  prediction <- data.table::as.data.table(
    DEGAS::predClassBag.optimized(
      ccModel = model,
      Exp = t_sc_mat,
      scORpat = "pat"
    )
  )

  data.table::setnames(
    prediction,
    old = names(prediction),
    new = colnames(pheno_df)
  )

  prediction[, cell_id := rownames(t_sc_mat)]

  prediction <- DEGAS::LabelBinaryCells(
    pred_dt = prediction,
    pheno_colnames = colnames(pheno_df),
    select_fraction = p$select_fraction,
    test_method = p$normality_test_method,
    min_threshold = p$min_thresh,
    verbose = verbose
  )

  metadata <- prediction[
    ,
    !names(prediction) %in% c("cell_id", "diff"),
    with = FALSE
  ]

  data.table::setnames(
    metadata,
    old = names(metadata),
    new = ifelse(
      names(metadata) == "label",
      "DEGAS",
      paste0("DEGAS_", names(metadata))
    )
  )

  metadata <- as.data.frame(metadata)
  rownames(metadata) <- prediction$cell_id

  sc_out <- SeuratObject::AddMetaData(
    sc_data,
    metadata = metadata
  )

  sc_out <- SigBridgeRUtils::AddMisc(
    seurat_obj = sc_out,
    DEGAS_type = label_type,
    DEGAS_para = d,
    cover = FALSE
  )

  list(
    scRNA_data = sc_out,
    model = model,
    DEGAS_prediction = prediction
  )
}

screen_DEGAS_uv_core <- function(
    sc_data,
    matched_bulk,
    phenotype,
    label_type,
    phenotype_class,
    verbose = TRUE,
    ...) {

  out <- run_DEGAS_uv_backend(
    sc_data = sc_data,
    matched_bulk = matched_bulk,
    phenotype = phenotype,
    label_type = label_type,
    phenotype_class = phenotype_class,
    verbose = verbose
  )

  list(
    scRNA_data = out$scRNA_data,
    result = out
  )
}
```

## 2.3 Validate and register the four executors

The current function is `ValidateScreenFunc()`, not `ScreenValFunc()`.

```{r}
#| label: validate-register

executors <- list(
  scPP_core = screen_scPP_core,
  LP_SGL_core = screen_LP_SGL_core,
  DEGAS_uv_core = screen_DEGAS_uv_core,
  Scissor_core = screen_Scissor_core
)

invisible(
  lapply(executors, SigBridgeR::ValidateScreenFunc)
)

SigBridgeR::Register(
  scPP_core = screen_scPP_core,
  supported_phenotypes = "binary",
  overwrite = TRUE,
  registry = "ScreenStrategy"
)

SigBridgeR::Register(
  LP_SGL_core = screen_LP_SGL_core,
  supported_phenotypes = "binary",
  overwrite = TRUE,
  registry = "ScreenStrategy"
)

SigBridgeR::Register(
  DEGAS_uv_core = screen_DEGAS_uv_core,
  supported_phenotypes = "binary",
  overwrite = TRUE,
  registry = "ScreenStrategy"
)

SigBridgeR::Register(
  Scissor_core = screen_Scissor_core,
  supported_phenotypes = "binary",
  overwrite = TRUE,
  registry = "ScreenStrategy"
)

stopifnot(all(names(executors) %in% names(SigBridgeR::ScreenStrategy)))
```

## 2.4 Run the four screens

The three R-oriented methods are dispatched concurrently. Each worker receives
a share of the half-core budget. DEGAS then runs in the main R process to avoid
forking or serialising a TensorFlow/reticulate session and receives the whole
half-core budget.

```{r}
#| label: run-screens

r_methods <- c(
  scPP = "scPP_core",
  LP_SGL = "LP_SGL_core",
  Scissor = "Scissor_core"
)

n_r_workers <- CFG$n_r_workers
threads_per_worker <- CFG$threads_per_worker

register_executor_in_worker <- function(method_name) {
  register_args <- list(executors[[method_name]])
  names(register_args) <- method_name

  do.call(
    SigBridgeR::Register,
    c(
      register_args,
      list(
        supported_phenotypes = "binary",
        overwrite = TRUE,
        registry = "ScreenStrategy",
        verbose = FALSE
      )
    )
  )
}

run_registered_screen <- function(method_name) {
  set.seed(CFG$seed)

  # Multisession workers have independent package namespaces and registries.
  register_executor_in_worker(method_name)

  # Avoid nested future backends. Parallelism occurs across methods here.
  future::plan(future::sequential)

  SigBridgeR::setThreads(
    threads = threads_per_worker,
    verbose = FALSE
  )

  SigBridgeR::Screen(
    matched_bulk = matched_bulk,
    sc_data = sc_data,
    phenotype = phenotype,
    label_type = CFG$phenotype_name,
    phenotype_class = CFG$phenotype_class,
    screen_method = method_name,
    verbose = TRUE
  )
}

future::plan(
  future::multisession,
  workers = n_r_workers
)

r_results <- future.apply::future_lapply(
  X = unname(r_methods),
  FUN = run_registered_screen,
  future.seed = CFG$seed
)

future::plan(future::sequential)
names(r_results) <- names(r_methods)

# TensorFlow/DEGAS is run after the multisession workers have closed.
SigBridgeR::setThreads(
  threads = CFG$cpu_budget,
  tf_config = list(
    inter_op = max(1L, floor(CFG$cpu_budget / 4L)),
    intra_op = CFG$cpu_budget
  ),
  verbose = TRUE
)

set.seed(CFG$seed)

degas_result <- SigBridgeR::Screen(
  matched_bulk = matched_bulk,
  sc_data = sc_data,
  phenotype = phenotype,
  label_type = CFG$phenotype_name,
  phenotype_class = CFG$phenotype_class,
  screen_method = "DEGAS_uv_core",
  verbose = TRUE
)

screen_results <- list(
  scPP = r_results$scPP,
  LP_SGL = r_results$LP_SGL,
  DEGAS = degas_result,
  Scissor = r_results$Scissor
)

saveRDS(screen_results, PATHS$screen_results)
```

## 2.5 Screening QC and saved method outputs

```{r}
#| label: screen-qc

METHOD_COLUMNS <- c(
  scPP = "ScPP",
  LP_SGL = "LP_SGL",
  DEGAS = "DEGAS",
  Scissor = "scissor"
)

screen_qc <- purrr::imap_dfr(
  METHOD_COLUMNS,
  function(column, method) {
    object <- screen_results[[method]]$scRNA_data

    if (!column %in% colnames(object[[]])) {
      stop(method, " did not create expected metadata column `", column, "`.")
    }

    as.data.frame(table(object[[]][[column]], useNA = "ifany")) |>
      tibble::as_tibble() |>
      dplyr::rename(status = Var1, n_cells = Freq) |>
      dplyr::mutate(method = method, .before = 1)
  }
)

readr::write_csv(
  screen_qc,
  file.path(PATHS$tables_dir, "screening_cell_counts.csv")
)

knitr::kable(screen_qc, caption = "Screening cell counts")
```

```{r}
#| label: save-lpsgl-deg

lpsgl_dge <- screen_results$LP_SGL$result$dge_res

if (!is.null(lpsgl_dge)) {
  readr::write_csv(
    as.data.frame(lpsgl_dge),
    file.path(PATHS$tables_dir, "LP_SGL_bulk_DEG.csv")
  )
}
```

```{r}
#| label: scissor-qc

scissor_reliability <- screen_results$Scissor$result$reliability_result
scissor_cell_evaluation <- screen_results$Scissor$result$cell_evaluation

if (!is.null(scissor_reliability)) {
  saveRDS(
    scissor_reliability,
    file.path(PATHS$objects_dir, "Scissor_reliability.rds")
  )
}

if (!is.null(scissor_cell_evaluation)) {
  saveRDS(
    scissor_cell_evaluation,
    file.path(PATHS$objects_dir, "Scissor_cell_evaluation.rds")
  )
}
```

# 3. Merge, weighted vote, and visualisation

## 3.1 Merge the four results

```{r}
#| label: merge-results

merged <- SigBridgeR::MergeResult(
  screen_results$scPP,
  screen_results$LP_SGL,
  screen_results$DEGAS,
  screen_results$Scissor
)

stopifnot(all(unname(METHOD_COLUMNS) %in% colnames(merged[[]])))
```

## 3.2 Harmonise labels and calculate the equally weighted vote

DEGAS can return `Other`; it is mapped explicitly to `Neutral` before voting
so all four methods use the same three labels.

```{r}
#| label: weighted-vote

normalise_screen_status <- function(x) {
  x <- as.character(x)
  x[is.na(x) | x == "" | x == "Other"] <- "Neutral"

  invalid <- setdiff(unique(x), c("Positive", "Negative", "Neutral"))
  if (length(invalid) > 0L) {
    stop(
      "Unexpected screening labels: ",
      paste(invalid, collapse = ", ")
    )
  }

  factor(
    x,
    levels = c("Negative", "Neutral", "Positive")
  )
}

for (column in unname(METHOD_COLUMNS)) {
  merged[[column]] <- normalise_screen_status(merged[[]][[column]])
}

vote_data <- merged[[]][
  ,
  unname(METHOD_COLUMNS),
  drop = FALSE
]
colnames(vote_data) <- names(METHOD_COLUMNS)

set.seed(CFG$seed)

merged$WeightedVote <- SigBridgeR::WeightedVote(
  data = vote_data,
  weights = SCREEN$consensus$method_weights,
  ties.method = SCREEN$consensus$ties.method
)

merged$WeightedVote <- normalise_screen_status(merged$WeightedVote)

vote_counts <- as.data.frame(
  table(merged$WeightedVote, useNA = "ifany")
) |>
  tibble::as_tibble() |>
  dplyr::rename(status = Var1, n_cells = Freq)

knitr::kable(vote_counts, caption = "Equally weighted consensus labels")

saveRDS(merged, PATHS$merged_result)
```

## 3.3 UpSet plot of positive-cell intersections

```{r}
#| label: upset-plot
#| fig-width: 12
#| fig-height: 7

upset_result <- SigBridgeR::ScreenUpset(
  screened_seurat = merged,
  screen_type = unname(METHOD_COLUMNS),
  order_by = c("freq", "degree"),
  n_intersections = 20L,
  show_plot = TRUE,
  verbose = TRUE
)

ggplot2::ggsave(
  filename = file.path(
    PATHS$figures_dir,
    "positive_cell_intersections_upset.pdf"
  ),
  plot = upset_result$plot,
  width = 12,
  height = 7
)

upset_stats <- upset_result$stats |>
  dplyr::mutate(
    sets = vapply(
      sets,
      function(x) paste(x, collapse = " & "),
      character(1)
    )
  )

readr::write_csv(
  upset_stats,
  file.path(PATHS$tables_dir, "positive_cell_intersections.csv")
)
```

## 3.4 Stacked bar plots by original cell type/state

```{r}
#| label: stacked-bars
#| fig-width: 14
#| fig-height: 10

fraction_result <- SigBridgeR::ScreenFractionPlot(
  screened_seurat = merged,
  group_by = COL$cell_annotation,
  screen_type = c(
    unname(METHOD_COLUMNS),
    "WeightedVote"
  ),
  show_null = TRUE,
  show_plot = TRUE,
  plot_title = "Phenotype-guided cell fractions",
  ncol = 2L,
  verbose = TRUE
)

fraction_plot <- if (!is.null(fraction_result$combined_plot)) {
  fraction_result$combined_plot
} else if (is.list(fraction_result$plot)) {
  patchwork::wrap_plots(fraction_result$plot, ncol = 2L)
} else {
  fraction_result$plot
}

print(fraction_plot)

ggplot2::ggsave(
  filename = file.path(
    PATHS$figures_dir,
    "screening_fractions_by_luque_annotation.pdf"
  ),
  plot = fraction_plot,
  width = 14,
  height = 10
)
```

## 3.5 One-page UMAP: original annotation plus the four methods

```{r}
#| label: umap-five-panel
#| fig-width: 18
#| fig-height: 10

status_colours <- c(
  Negative = "#386C9B",
  Neutral = "#CECECE",
  Positive = "#FF3333"
)

p_annotation <- Seurat::DimPlot(
  merged,
  reduction = "umap",
  group.by = COL$cell_annotation,
  label = TRUE,
  repel = TRUE,
  raster = TRUE
) +
  ggplot2::ggtitle("Original cell type/state") +
  Seurat::NoLegend()

method_plots <- purrr::imap(
  METHOD_COLUMNS,
  function(column, method) {
    Seurat::DimPlot(
      merged,
      reduction = "umap",
      group.by = column,
      cols = status_colours,
      raster = TRUE
    ) +
      ggplot2::ggtitle(method)
  }
)

umap_page <- patchwork::wrap_plots(
  c(list(p_annotation), method_plots),
  ncol = 3L,
  guides = "collect"
) +
  patchwork::plot_annotation(
    title = "Neural-biased morphotype screening",
    subtitle = paste0(
      "Original `", COL$cell_annotation,
      "` plus Neutral, Positive, and Negative calls"
    )
  )

print(umap_page)

ggplot2::ggsave(
  filename = file.path(
    PATHS$figures_dir,
    "umap_original_and_four_methods.pdf"
  ),
  plot = umap_page,
  width = 18,
  height = 10
)
```

## 3.6 Optional consensus UMAP and final output inventory

```{r}
#| label: consensus-umap
#| fig-width: 8
#| fig-height: 7

p_consensus <- Seurat::DimPlot(
  merged,
  reduction = "umap",
  group.by = "WeightedVote",
  cols = status_colours,
  raster = TRUE
) +
  ggplot2::ggtitle("Equally weighted consensus")

print(p_consensus)

ggplot2::ggsave(
  filename = file.path(
    PATHS$figures_dir,
    "umap_weighted_vote.pdf"
  ),
  plot = p_consensus,
  width = 8,
  height = 7
)
```

```{r}
#| label: output-inventory

output_inventory <- tibble::tibble(
  output = c(
    "Preprocessed single-cell object",
    "Filtered bulk counts",
    "Log-CPM bulk screening matrix",
    "Validated phenotype",
    "Four method results",
    "Merged object with weighted vote",
    "UpSet plot",
    "Stacked-bar plots",
    "Five-panel UMAP"
  ),
  path = c(
    PATHS$preprocessed_sc,
    PATHS$preprocessed_bulk_counts,
    PATHS$preprocessed_bulk_expression,
    PATHS$phenotype,
    PATHS$screen_results,
    PATHS$merged_result,
    file.path(PATHS$figures_dir, "positive_cell_intersections_upset.pdf"),
    file.path(
      PATHS$figures_dir,
      "screening_fractions_by_luque_annotation.pdf"
    ),
    file.path(
      PATHS$figures_dir,
      "umap_original_and_four_methods.pdf"
    )
  )
)

knitr::kable(output_inventory, caption = "Pipeline outputs")
```

# Reproducibility record

```{r}
#| label: session-info

sessionInfo()
```

# Documentation used

- [SigBridgeR full tutorial](https://wanglabcsu.github.io/SigBridgeR/articles/Full_Tutorial.html)
- [SigBridgeR deep-learning overview](https://deepwiki.com/WangLabCSU/SigBridgeR/3.3-deep-learning-methods-(degas-sidish))
- [Current SigBridgeR reference index](https://wanglabcsu.github.io/SigBridgeR/reference/index.html)
- [SigBridgeR source repository](https://github.com/WangLabCSU/SigBridgeR)
